How has the Brazilian Amazon been constructed as a problem? Presidential speeches and transnational politics since 1985

### 2022-09-03 Draft_September_2022/Full_draft_20220901.rmd
More updates to article

---
title: How has the Brazilian Amazon been constructed as a problem? Presidential speeches and transnational politics since 1985 in Brazil
author:
- Livio Silva-Muller
- Henrique Sposito^[Both co-authors contributed equally to the article. Names are ordered alphabetically. Livio is a PhD Candidate in Anthropology and Sociology at the Geneva Graduate Institute and affiliated to the Albert Hirschman Centre on Democracy. Henrique is a PhD Candidate in International Relations and Political Sciences at the Geneva Graduate Institute and affiliated to the Centre for International Environmental Studies.]
date: "1 September 2022"
output:
  pdf_document: default
  latex_engine: xelatex
  fig_caption: yes
  html_document:
    df_print: paged
fontsize: 12pt
mainfont: Times New Roman
header-includes:
   - \usepackage{floatrow}
   - \floatsetup[figure]{capposition=top}
   - \usepackage{float}
   - \floatplacement{figure}{H}
   - \usepackage[T1]{fontenc}
   - \usepackage[utf8]{inputenc}
   - |
    ```{=latex}
    \providecommand{\keywords}[1]{\textbf{Keywords:} #1}
    \providecommand{\wordcount}[1]{\text{Word Count:}#1} 
    ```
always_allow_html: TRUE
bibliography: references.bib
csl: chicago.csl
abstract: Presidential speeches influence the ways we think about, and act towards, the environment transnationally. Still, these remain understudied in this context. We propose a framework to investigate how the Brazilian Amazon has been constructed as a problem in discourses along time and space. Using supervised machine learning, we classify statements about the Amazon in 6130 presidential speeches since 1985. We find that national and international events drive the frequency of the Amazon as a topic in presidential speeches inside and outside Brazil. While constructing the Amazon a problem of economic integration dominates discourses until the mid-2000s, environmental conservation and social development constructions increase in relevance and temporarily surpass economic integration from 2010 to 2015. In turn, constructing the Amazon as an issue of sovereignty steadily increases since 2010. Lastly, the farthest presidents are from the Amazon, the more likely they are to construct the Amazon as an issue of environmental conservation.
---

```{=latex}
\keywords{ discourse analysis, transnational governance, environmental policy, Brazilian Amazon, supervised learning, and deforestation}
```

@@

When, where, and how often do presidents speak about the Amazon? Figure 2, below, shows the proportion of speeches that mentions the Amazon in relation to all speeches in a year. On average, presidents mention the Amazon in 15% ^[For comparison purposes the averages for other policy objects in the same corpus of speeches are: Inequality (14.7%), Criminality (17.3%), Inflation (20.9%), and Unemployment (13.8%).] of their speeches . Mentions of the Amazon outside of Brazil are slightly more frequent than mentions inside Brazil (Figure 3). The frequency in which the Amazon appears across presidential speeches inside, outside, and overall, though, varies similarly across time.

@@

1990-91 the Amazon appeared in 12% to 14% of all speeches but, by 1992, this average increased to 25%. The 1992 Earth Summit in Brazil brought international attention to environmental topics related to the Amazon. One of the big announcements, for example, was the consolidation of the first transnational partnership for the Amazon, the G7 Pilot Program, which brought a high number of financial resources to the region for public policy implementation [@capobianco2021]. Throughout the rest of the 1990s and 2000s, during the Cardoso and Lula administrations, the Amazon as a topic did not diverge much from the average until 2009, when we see another spike in mentions of the Amazon. This coincides with the 2009 Copenhagen Summit and the steepest decrease in deforestation rates since data is available as Figure 5, below, illustrates. Lula led the delegation to Copenhagen with a self-image of “we do not promise, we deliver” when the stakes about climate change were high [@franchini2019].

```{r Figure 3, fig.cap="Amazonian speeches in time and space", echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE, include=TRUE}
perc_in <- pop %>%
   mutate(location_cat = ifelse(grepl("International|Non Identified",
                                      location_cat),
                                "Outside of Brazil", "In Brazil")) %>%
  filter(location_cat == "In Brazil") %>% 
  count(amazon_speech, year, location_cat) %>%
  group_by(year) %>%
  mutate(perc = n / sum(n)) %>%
  filter(amazon_speech == 1) %>%
  select(c(perc, amazon_speech, year, location_cat))
perc_out <- pop %>%
   mutate(location_cat = ifelse(grepl("International|Non Identified",
                                      location_cat),
                                "Outside of Brazil", "In Brazil")) %>%
  filter(location_cat == "Outside of Brazil") %>% 
  count(amazon_speech, year, location_cat) %>%
  group_by(year) %>%
  mutate(perc = n / sum(n)) %>%
  filter(amazon_speech == 1) %>%
  select(c(perc, amazon_speech, year, location_cat)) 
# location plots
Amazon_year_in <- perc_in %>%
  ggplot(., aes(x = year, y = perc))+
  geom_line(color="darkgreen")+
  geom_line(aes(y=mean(perc)), color="black", size=.5, linetype="dotted")+
  annotate("text", x = 1985, y = 0.17, label = "16.17%", size = 3.5)+
  annotate("text", x = 1988.7, y = 0.35, label = "Chico Mendes", size = 3)+
  annotate("text", x = 1992, y = 0.35, label = "Earth Summit", size = 3)+
  annotate("text", x = 2009, y = 0.23, label = "Copenhagen Summit", size = 3)+
  annotate("text", x = 2015, y = 0.20, label = "Paris Summit", size = 3)+
  annotate("text", x = 2019, y = 0.25, label = "August Fires", size=3) +
  scale_y_continuous(labels = percent_format())+
  labs(x = "", y = "",
       title = "Share of speeches inside Brazil",
       subtitle= " 4344 speeches out of 6130",
       color='Y series') +
  theme_am()
Amazon_year_out <- perc_out %>% 
  ggplot(., aes(x = year, y = perc))+
  geom_line(color="darkgreen")+
  geom_line(aes(y=mean(perc)), color="black", size=.5, linetype="dotted")+
  annotate("text", x = 1985, y = 0.165, label = "15.49%", size = 3.5)+  
  annotate("text", x = 1988.7, y = 0.33, label = "Chico Mendes", size = 3)+
  annotate("text", x = 1992, y = 0.33, label = "Earth Summit", size = 3)+
  annotate("text", x = 2009, y = 0.25, label = "Copenhagen Summit", size = 3)+
  annotate("text", x = 2015, y = 0.19, label = "Paris Summit", size = 3)+
  annotate("text", x = 2019, y = 0.24, label = "August Fires", size=3) +
  scale_y_continuous(labels = percent_format())+
  labs(x = "", y = "",
 title = "Share of speeches outside Brazil",
       subtitle= " 1786 speeches out of 6130",
       color='Y series') +
  theme_am()
grid.arrange(Amazon_year_in, Amazon_year_out, nrow=2)
```

@@

The trends identified in Figure 4 are especially interesting considering policies related to deforestation in the Amazon and deforestation rates portrayed in Figure 5, below. First, deforestation rates have a strong negative correlation with environmental conservation. When deforestation rates increased from 1988 to 2004, and from 2012 until 2021, environmental conservation constructions decreased. As deforestation rates decreased from 2004 to 2012, environmental conservation constructions in presidential speeches increased. @capobianco2021 argues that the unprecedented decrease in deforestation we observed from 2004 to 2012 was a product of an increase in the perception of stronger federal policies and police presence in the Amazon region which, in turn, engendered a perception of a higher risk of being caught and fined for deforestation. The shift from constructing the Amazon as an issue of economic integration to an issue of environmental conservation generates a perception that illegal deforestation will be increasingly monitored and sanctioned. Second and relatedly, deforestation rates have a positive and strong correlation with economic integration. When deforestation increased from 1988 to 2004 and from 2012 to 2021, so did economic integration. When deforestation decreased from 2004 to 2012, economic integration decreased as well. This suggests that presidents might boast policy outcomes related to economic integration as justifications for deforestation.

```{r Figure 5, fig.cap="Deforestation rates since 1988", echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, include=TRUE}
def_rates <- ggplot(amazon_def_year, aes(x = year)) +
  geom_line(aes(y = area), color= "darkolivegreen", linetype="dotted", size= 1.5)+
  geom_smooth(se=FALSE, color="black", aes(y=area), linetype="solid", size= 0.75)+
  annotate("text", x = 2004, y = 29000,
           label = "2004 Plan to Combatand Control Deforestation", size = 3) +
  annotate("text", x = 2012.5, y = 3500, label = "2012 Forest Code ", size = 3) +
  labs(x = "", y = "Deforestation in squared kilometers",
       title = "Deforestation rates since 1988",
       subtitle= "Data from INPE, Brazilian Ministry of Science and Technology.",
       color = "",
       caption= "Black curve smoothed using loess.")+
  theme_am()
def_rates
```

@@

At the level of the speech presidents might mix multiple problem constructions. Although presidents prefer pure problem-constructions, they also construct the Amazon as a multifaceted issue. Mixed-type problem constructions in discourse offer more intricate understandings of the Amazon as a problem. Constructing the Amazon as multiple issues averages at 18% of all constructions in time. Some mixed-types rarely appear so we focus our discussion on those mixed-types with higher incidence. Figure 6, below, displays the averages of mixed-types constructions by presidents.

```{r Figure 6, fig.cap="Mixed-types by president", echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, include=TRUE}
mixed_types_presid <- ama_mx %>%
  count(president, mx_cat, mx_cat2) %>%
  group_by(president) %>%
  mutate(perc = n / sum(n)) %>%
  filter(mx_cat2 == "Mixed-types") %>%
  mutate(president = factor(president, levels = c("Sarney", "Collor", "Itamar", "FHC",
                                                  "Lula", "Dilma", "Temer", "Bolsonaro"))) %>%
  ggplot(aes(x = president, y = perc, fill = mx_cat,
             label = scales::percent(perc, accuracy=.1))) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(size = 2, position = position_stack(vjust = .5)) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  scale_fill_brewer(type = "qual", palette = "Set2", direction= 1) +
  labs(x = "", y = "", title ="Share of each mixed-type problem-constructions by president",
       subtitle = "Sample composed by 1895 Amazonian statements in presidential speeches since 1985",
       caption = "") +
  theme_am() +
  theme(legend.position = "right")
mixed_types_presid
```

The most frequent mixed-type problem construction of the Amazon is economic conservation. This mixed-type construction, composed of Amazonian statements that construct the Amazon as a problem of both economic integration and environmental conservation, generally increased in time. Its increase over time suggests that later presidents, along with being more diverse in how they constructed the amazon as an issue (see Figure 3), increasingly find ways to reconcile the dictum between the economy and the environment that prevailed in the previous years. The second most common mixed-type mixes social development and economic integration and appears, on average, in 5.4% of all statements. This is not surprising, as developing countries like Brazil repeatedly claim a “right to develop” when it comes to negotiating strong climate commitments and policies with a focus on both social development and economic integration. Nonetheless, there is a general increase of mixed-types in time, which is expected given the rise of global agendas understanding interconnections of social, environmental, and economic domains such as the Millennium Development Goals and the Sustainable Development Goals.

@@

The Amazon region, forests, and peoples have historically been the topic of local, national, and international debates. This implies that audiences’ priorities in each setting change and which policies are appropriate to solve the “Amazon issue” differ as well. As presidents move away from the Amazon, audiences might have different stakes and interests in it. While speeches inside the Amazon might focus on social development, for example, speeches in the United Nations in New York might highlight environmental conservation. To investigate the relationship between where presidents speak and how they construct the Amazon as an issue, we run four logistic regressions, one for each pure-type construction as a dependent variable. We focus on pure-type constructions as these represent most constructions and avoid ambiguity. The independent variable for all models is the distance between the state country capital to Manaus. Figure 7, below, depicts the predicted probabilities from the models (see Table 1) as functions of distance.

```{r Figure 7, fig.cap="Percent of problem constructions across locations", echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, include=TRUE}
ama_model <- dplyr::left_join(ama_mx, amazon_def_year, by = "year") # merge
ama_model <- dplyr::left_join(ama_model, AAI, by = "year") # merge
ama_model <- filter(ama_model, location_cat != "Non Identified",
                    mx_cat != "Other Mixed-types") %>%
  mutate(con_vs_all = ifelse(mx_cat == "Pure Conservation", 1, 0),
         EI_vs_all = ifelse(mx_cat == "Pure Economic Integration", 1, 0),
         SD_vs_all = ifelse(mx_cat == "Pure Social Development", 1, 0),
         sov_vs_all = ifelse(mx_cat == "Pure Sovereignty", 1, 0))
# model
model_logit_con <- glm(con_vs_all ~ km_to_manaus + election_year + def_year + AAI,
                   family=binomial(link = "logit"), data = ama_model)
model_logit_ei <- glm(EI_vs_all ~ km_to_manaus + election_year + def_year + AAI,
                   family=binomial(link = "logit"), data = ama_model)
model_logit_sd <- glm(SD_vs_all ~ km_to_manaus + election_year + def_year + AAI,
                   family=binomial(link = "logit"), data = ama_model)
model_logit_sov <- glm(sov_vs_all ~ km_to_manaus + election_year + def_year + AAI,
                   family=binomial(link = "logit"), data = ama_model)
# Plot model
ama_model$"Environmental Conservation" = predict(model_logit_con, ama_model, type = "response")
ama_model$"Economic Integration" = predict(model_logit_ei, ama_model, type = "response")
ama_model$"Social Development" = predict(model_logit_sd, ama_model, type = "response")
ama_model$"National Sovereignty" = predict(model_logit_sov, ama_model, type = "response")
plot_loc <- ama_model %>% gather(key=p_c, value = pred, 22:24)
plot_loc$p_c <-factor(plot_loc$p_c, levels = c("Economic Integration",
                                               "Environmental Conservation",
                                               "Social Development"))
all_loc <- ggplot(plot_loc, aes(x = km_to_manaus, y=pred, color=p_c)) +
  geom_jitter(alpha=1, size=1)+
  geom_smooth(size = .5, se=FALSE) +
  geom_vline(xintercept = c(1933.78, 2771.49, 5075.48, 8306.76), linetype = "dotted") +
  ggplot2::annotate(geom = "text", x = c(1933.78,2771.49 , 5075.48, 8306.76),
                    y = c(0.5,0.55,0.5,0.55),
                    label = c("Brasilia","Sao Paulo", "New York", "Paris"), alpha = 0.4) +
  scale_y_continuous(labels = percent_format()) + 
  labs(x = "Distance from the Amazon (km)",
       y = "",
       title = "Predicted share of each problem-construction as a function of distance from the Amazon",
       subtitle="Sovereignty was removed because of statistical insignificance (table 1).", 
       caption = "Curve in the plot is estimated using loess method.") +
    theme_am() +
  theme(legend.position = "bottom")
all_loc
```

@@

\begin{landscape}
```{r Table 1, fig.cap="Logistic Regressions Output", echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, include=TRUE, results="asis"}
# summary statistics
md <-stargazer::stargazer(model_logit_con, model_logit_ei, model_logit_sd,
                          model_logit_sov, digits = 10, 
                          dep.var.labels = c("Conservation", "Economic Integration",
                                             "Social Development", "Sovereignty"),
                          covariate.labels = c("Distance from the Amazon (km)",
                                               "Election year", "Yearly Deforestation",
                                               "Yearly   Inflation"), no.space = TRUE)
```

@@

>>>>

---
title: How has the Brazilian Amazon been constructed as a problem? Presidential speeches and transnational politics since 1985
author:
- Livio Silva-Muller
- Henrique Sposito^[Both co-authors contributed equally to the article. Names are ordered alphabetically. Livio is a PhD Candidate in Anthropology and Sociology at the Geneva Graduate Institute and affiliated to the Albert Hirschman Centre on Democracy. Henrique is a PhD Candidate in International Relations and Political Sciences at the Geneva Graduate Institute and affiliated to the Centre for International Environmental Studies.]
date: "1 September 2022"
output:
  pdf_document: default
  latex_engine: xelatex
  fig_caption: yes
  html_document:
    df_print: paged
fontsize: 12pt
mainfont: Times New Roman
header-includes:
   - \usepackage{floatrow}
   - \floatsetup[figure]{capposition=top}
   - \usepackage{float}
   - \floatplacement{figure}{H}
   - \usepackage[T1]{fontenc}
   - \usepackage[utf8]{inputenc}
   - |
    ```{=latex}
    \providecommand{\keywords}[1]{\textbf{Keywords:} #1}
    \providecommand{\wordcount}[1]{\text{Word Count:}#1} 
    ```
always_allow_html: TRUE
bibliography: references.bib
csl: chicago.csl
abstract: Presidential speeches influence the ways we think about, and act towards, the environment transnationally. Still, these remain understudied in this context. We propose a framework to investigate how the Brazilian Amazon has been constructed as a problem in discourses along time and space. Using supervised machine learning, we classify statements about the Amazon in 6130 presidential speeches since 1985. We find that national and international events drive the frequency of the Amazon as a topic in presidential speeches inside and outside Brazil. While constructing the Amazon a problem of economic integration dominates discourses until the mid-2000s, environmental conservation and social development constructions increase in relevance and temporarily surpass economic integration from 2010 to 2015. In turn, constructing the Amazon as an issue of sovereignty steadily increases since 2010. Lastly, the farthest presidents are from the Amazon, the more likely they are to construct the Amazon as an issue of environmental conservation.
---

```{=latex}
\keywords{ discourse analysis, transnational governance, environmental policy, Brazilian Amazon, supervised learning, deforestation}
```

@@

When, where, and how often do presidents speak about the Amazon? Figure 2, below, shows the proportion of speeches that mentions the Amazon in relation to all speeches in a year. On average, presidents mention the Amazon in 15% ^[For comparison purposes the averages for other policy objects in the same corpus of speeches are: Inequality (14.7%), Criminality (17.3%), Inflation (20.9%), and Unemployment (13.8%).] of their speeches. The frequency in which the Amazon appears across presidential speeches inside, outside, and overall varies similarly across time (see Figure 7 in appendix).

@@

1990-91 the Amazon appeared in 12% to 14% of all speeches but, by 1992, this average increased to 25%. The 1992 Earth Summit in Brazil brought international attention to environmental topics related to the Amazon. One of the big announcements, for example, was the consolidation of the first transnational partnership for the Amazon, the G7 Pilot Program, which brought a high number of financial resources to the region for public policy implementation [@capobianco2021]. Throughout the rest of the 1990s and 2000s, during the Cardoso and Lula administrations, the Amazon as a topic did not diverge much from the average until 2009, when we see another spike in mentions of the Amazon. This coincides with the 2009 Copenhagen Summit and the steepest decrease in deforestation rates since data is available as Figure 3, below, illustrates. Lula led the delegation to Copenhagen with a self-image of “we do not promise, we deliver” when the stakes about climate change were high [@franchini2019].

```{r Figure 3, fig.cap="Deforestation rates since 1988", echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, include=TRUE}
def_rates <- ggplot(amazon_def_year, aes(x = year)) +
  geom_line(aes(y = area), color= "darkolivegreen", linetype="dotted", size= 1.5)+
  geom_smooth(se=FALSE, color="black", aes(y=area), linetype="solid", size= 0.75)+
  annotate("text", x = 2004, y = 29000,
           label = "2004 Plan to Combatand Control Deforestation", size = 3) +
  annotate("text", x = 2012.5, y = 3500, label = "2012 Forest Code ", size = 3) +
  labs(x = "", y = "Deforestation in squared kilometers",
       title = "Deforestation rates since 1988",
       subtitle= "Data from INPE, Brazilian Ministry of Science and Technology.",
       color = "",
       caption= "Black curve smoothed using loess.")+
  theme_am()
def_rates
```

@@

The trends identified are especially interesting considering Amazonian policies and deforestation rates portrayed in Figure 3, above. First, deforestation rates have a strong negative correlation with environmental conservation. When deforestation rates increased from 1988 to 2004, and from 2012 until 2021, environmental conservation constructions decreased. As deforestation rates decreased from 2004 to 2012, environmental conservation constructions in presidential speeches increased. @capobianco2021 argues that the unprecedented decrease in deforestation we observed from 2004 to 2012 was a product of an increase in the perception of stronger federal policies and police presence in the Amazon region which, in turn, engendered a perception of a higher risk of being caught and fined for deforestation. The shift from constructing the Amazon as an issue of economic integration to an issue of environmental conservation generates a perception that illegal deforestation will be increasingly monitored and sanctioned. Second and relatedly, deforestation rates have a positive and strong correlation with economic integration. When deforestation increased from 1988 to 2004 and from 2012 to 2021, so did economic integration. When deforestation decreased from 2004 to 2012, economic integration decreased as well. This suggests that presidents might boast policy outcomes related to economic integration as justifications for deforestation.

@@

At the level of the speech presidents might mix multiple problem constructions. Although presidents prefer pure problem-constructions, they also construct the Amazon as a multifaceted issue. Mixed-type problem constructions in discourse offer more intricate understandings of the Amazon as a problem. Constructing the Amazon as multiple issues averages at 18% of all constructions in time. Some mixed-types rarely appear so we focus our discussion on those mixed-types with higher incidence. Figure 5, below, displays the averages of mixed-types constructions by presidents.

```{r Figure 5, fig.cap="Mixed-types by president", echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, include=TRUE}
mixed_types_presid <- ama_mx %>%
  count(president, mx_cat, mx_cat2) %>%
  group_by(president) %>%
  mutate(perc = n / sum(n)) %>%
  filter(mx_cat2 == "Mixed-types") %>%
  mutate(president = factor(president, levels = c("Sarney", "Collor", "Itamar", "FHC",
                                                  "Lula", "Dilma", "Temer", "Bolsonaro"))) %>%
  ggplot(aes(x = president, y = perc, fill = mx_cat,
             label = scales::percent(perc, accuracy=.1))) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(size = 2, position = position_stack(vjust = .5)) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  scale_fill_brewer(type = "qual", palette = "Set2", direction= 1) +
  labs(x = "", y = "", title ="Share of each mixed-type problem-constructions by president",
       subtitle = "Sample composed by 1895 Amazonian statements in presidential speeches since 1985",
       caption = "") +
  theme_am() +
  theme(legend.position = "right")
mixed_types_presid
```

The most frequent mixed-type problem construction of the Amazon is economic conservation. This mixed-type construction, composed of Amazonian statements that construct the Amazon as a problem of both economic integration and environmental conservation, generally increased in time. Its increase over time suggests that later presidents, along with being more diverse in how they constructed the amazon as an issue (see Figure 4), increasingly find ways to reconcile the dictum between the economy and the environment that prevailed in the previous years. The second most common mixed-type mixes social development and economic integration and appears, on average, in 5.4% of all statements. This is not surprising, as developing countries like Brazil repeatedly claim a “right to develop” when it comes to negotiating strong climate commitments and policies with a focus on both social development and economic integration. Nonetheless, there is a general increase of mixed-types in time, which is expected given the rise of global agendas understanding interconnections of social, environmental, and economic domains such as the Millennium Development Goals and the Sustainable Development Goals.

@@

The Amazon region, forests, and peoples have historically been the topic of local, national, and international debates. This implies that audiences’ priorities in each setting change and which policies are appropriate to solve the “Amazon issue” differ as well. As presidents move away from the Amazon, audiences might have different stakes and interests in it. While speeches inside the Amazon might focus on social development, for example, speeches in the United Nations in New York might highlight environmental conservation. To investigate the relationship between where presidents speak and how they construct the Amazon as an issue, we run four logistic regressions, one for each pure-type construction as a dependent variable. We focus on pure-type constructions as these represent most constructions and avoid ambiguity. The independent variable for all models is the distance between the state country capital to Manaus. Figure 6, below, depicts the predicted probabilities from the models (see Table 1) as functions of distance.

```{r Figure 6, fig.cap="Percent of problem constructions across locations", echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, include=TRUE}
ama_model <- dplyr::left_join(ama_mx, amazon_def_year, by = "year") # merge
ama_model <- dplyr::left_join(ama_model, AAI, by = "year") # merge
ama_model <- filter(ama_model, location_cat != "Non Identified",
                    mx_cat != "Other Mixed-types") %>%
  mutate(con_vs_all = ifelse(mx_cat == "Pure Conservation", 1, 0),
         EI_vs_all = ifelse(mx_cat == "Pure Economic Integration", 1, 0),
         SD_vs_all = ifelse(mx_cat == "Pure Social Development", 1, 0),
         sov_vs_all = ifelse(mx_cat == "Pure Sovereignty", 1, 0))
# model
model_logit_con <- glm(con_vs_all ~ km_to_manaus + election_year + def_year + AAI,
                   family=binomial(link = "logit"), data = ama_model)
model_logit_ei <- glm(EI_vs_all ~ km_to_manaus + election_year + def_year + AAI,
                   family=binomial(link = "logit"), data = ama_model)
model_logit_sd <- glm(SD_vs_all ~ km_to_manaus + election_year + def_year + AAI,
                   family=binomial(link = "logit"), data = ama_model)
model_logit_sov <- glm(sov_vs_all ~ km_to_manaus + election_year + def_year + AAI,
                   family=binomial(link = "logit"), data = ama_model)
# Plot model
ama_model$"Environmental Conservation" = predict(model_logit_con, ama_model, type = "response")
ama_model$"Economic Integration" = predict(model_logit_ei, ama_model, type = "response")
ama_model$"Social Development" = predict(model_logit_sd, ama_model, type = "response")
ama_model$"National Sovereignty" = predict(model_logit_sov, ama_model, type = "response")
plot_loc <- ama_model %>% gather(key=p_c, value = pred, 22:24)
plot_loc$p_c <-factor(plot_loc$p_c, levels = c("Economic Integration",
                                               "Environmental Conservation",
                                               "Social Development"))
all_loc <- ggplot(plot_loc, aes(x = km_to_manaus, y=pred, color=p_c)) +
  geom_jitter(alpha=1, size=1) +
  geom_smooth(size = .5, se=FALSE) +
  geom_vline(xintercept = c(1933.78, 2771.49, 5075.48, 8306.76), linetype = "dotted") +
  ggplot2::annotate(geom = "text", x = c(1933.78,2771.49 , 5075.48, 8306.76),
                    y = c(0.45,0.5,0.47,0.5),
                    label = c("Brasilia", "Sao Paulo", "New York", "Paris"), alpha = 0.4) +
  scale_y_continuous(labels = percent_format()) +
  labs(x = "Distance from the Amazon (km)",
       y = "",
       title = "Predicted share of each problem-construction as a function of distance from the Amazon",
       subtitle="Sovereignty was removed because of statistical insignificance (table 1).", 
       caption = "Curve in the plot is estimated using loess method.") +
    theme_am() +
  theme(legend.position = "bottom")
all_loc
```

@@

\begin{landscape}

```{r Table 1, fig.cap="Logistic Regressions Output", echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, include=TRUE, results="asis"}
# summary statistics
md <-stargazer::stargazer(model_logit_con, model_logit_ei, model_logit_sd,
                          model_logit_sov, digits = 10, 
                          dep.var.labels = c("Conservation", "Economic Integration",
                                             "Social Development", "Sovereignty"),
                          covariate.labels = c("Distance from the Amazon (km)",
                                               "Election year", "Yearly Deforestation",
                                               "Yearly   Inflation"), no.space = TRUE)
```

@@

```{r Figure 7, fig.cap="Amazonian speeches in time and space", echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE, include=TRUE}
perc_in <- pop %>%
   mutate(location_cat = ifelse(grepl("International|Non Identified",
                                      location_cat),
                                "Outside of Brazil", "In Brazil")) %>%
  filter(location_cat == "In Brazil") %>% 
  count(amazon_speech, year, location_cat) %>%
  group_by(year) %>%
  mutate(perc = n / sum(n)) %>%
  filter(amazon_speech == 1) %>%
  select(c(perc, amazon_speech, year, location_cat))
perc_out <- pop %>%
   mutate(location_cat = ifelse(grepl("International|Non Identified",
                                      location_cat),
                                "Outside of Brazil", "In Brazil")) %>%
  filter(location_cat == "Outside of Brazil") %>% 
  count(amazon_speech, year, location_cat) %>%
  group_by(year) %>%
  mutate(perc = n / sum(n)) %>%
  filter(amazon_speech == 1) %>%
  select(c(perc, amazon_speech, year, location_cat)) 
# location plots
Amazon_year_in <- perc_in %>%
  ggplot(., aes(x = year, y = perc))+
  geom_line(color="darkgreen")+
  geom_line(aes(y=mean(perc)), color="black", size=.5, linetype="dotted")+
  annotate("text", x = 1985, y = 0.17, label = "16.17%", size = 3.5)+
  annotate("text", x = 1988.7, y = 0.39, label = "Chico Mendes", size = 3)+
  annotate("text", x = 1992, y = 0.36, label = "Earth Summit", size = 3)+
  annotate("text", x = 2009, y = 0.28, label = "Copenhagen Summit", size = 3)+
  annotate("text", x = 2015, y = 0.24, label = "Paris Summit", size = 3)+
  annotate("text", x = 2019, y = 0.28, label = "August Fires", size=3) +
  scale_y_continuous(labels = percent_format())+
  labs(x = "", y = "",
       title = "Share of speeches inside Brazil",
       subtitle= " 4344 speeches out of 6130",
       color='Y series') +
  theme_am()
Amazon_year_out <- perc_out %>% 
  ggplot(., aes(x = year, y = perc))+
  geom_line(color="darkgreen")+
  geom_line(aes(y=mean(perc)), color="black", size=.5, linetype="dotted")+
  annotate("text", x = 1985, y = 0.165, label = "15.49%", size = 3.5)+  
  annotate("text", x = 1988.7, y = 0.39, label = "Chico Mendes", size = 3)+
  annotate("text", x = 1992, y = 0.36, label = "Earth Summit", size = 3)+
  annotate("text", x = 2009, y = 0.28, label = "Copenhagen Summit", size = 3)+
  annotate("text", x = 2015, y = 0.24, label = "Paris Summit", size = 3)+
  annotate("text", x = 2019, y = 0.28, label = "August Fires", size=3)+
  scale_y_continuous(labels = percent_format())+
  labs(x = "", y = "",
 title = "Share of speeches outside Brazil",
       subtitle= " 1786 speeches out of 6130",
       color='Y series') +
  theme_am()
grid.arrange(Amazon_year_in, Amazon_year_out, nrow=2)
```
